Semantic browsing of sound databases \\ without keywords

### 2016-07-05 lafaySsf2016.tex
Derniere relecture, Mathias RossignolL

We thus consider in this paper three means of displaying sounds without relying on any textual representation, all based on a spatial organization of sounds. The first one, considered as a reference baseline, does not rely on any semantic information and positions sounds according to their acoustical properties (time averaged spectral features). The second and third displays exhibit a spatial organization based  on a predefined hierarchical semantic organization; those last two displays differs by how semantics are used.

@@

As a reference, an Acoustic Display (AD) provides a spatial representation based on acoustic descriptors. Each sound is described by time-averaged Mel-Frequency Cepstrum Coefficients (MFCCs) computed with standard parameter settings (13 lowest quefrency coefficients are kept out of 40). The Euclidean distance between time averaged features is then computed for each sound pair. A non metric multidimensional scaling (MDS) with Kruskal's normalized stress \cite{kruskal1964multidimensional} is then employed to project the data into two dimensions according to this distance, as shown on Figure~\ref{figXP3music}. The latter exhibit a clear distinction between percussive (left) and sustained sounds (right).

@@

Among the several alternatives that exists, MFCCs features are chosen as they are widely used in the audio processing community. MDS is selected for 2D projection as it has a lower sensitivity to outliers than the PCA and do not relyi on any parameter optimization, unlike the SOM approach.

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{gfx/music_mds.png} 
\end{center}
\caption{\label{figXP3music} Acoustical Display for the musical sound dataset (AD). Colors and legend are removed during the experiment.}
\end{figure}

Two semantically oriented displays are then proposed, called Full Semantic Display (FSD) and Progressive Semantic Display (PSD). Both display  consider the hierarchical structure of the dataset to organize sounds. Each sound class is represented by a filled circle. Circles are then packed together according to the hierarchical semantic organization of the dataset, as shown in Figure~\ref{figSSF}. Thus, subclasses belonging to the same class are close to each others. Circle packing functions of the D3.js (Data-Driven Documents) javascript library \cite{2011-d3} are used to distribute the sound classes in the space. Depending on the display, the user can either access each leaf class directly (FSD), or has to click through intermediate levels of the hierarchy (PSD).

@@

During this experiment, the three displays are compared using two corpora: the urban sound corpus and the musical sound corpus. By comparing AD and xSD (the two semantic displays), the goal is to evaluate the  relative efficiencies of semantic based and acoustic based spatial configurations. By comparing FSD and PSD, we study the impact of an enforced hierarchical exploration of the corpus by the user. The goal is to check if constraining the user to first browse the high levels of the hierarchy helps him to grasp and memorize the spatial configuration and the organization of the sound classes.

@@

We evaluate and compare those three displays with a crowdsourcing web-based experiment\footnote{The tests are available at \url{http://soundthings.org/research/speedSoundFinding/index.html} and \url{http://soundthings.org/research/speedSoundFindingMusic/index.html}, each of the 3 displays is also available by appending a digit from 1 to 3 to \textit{index} in the above cited url.} for which we adopt a between subject approach, \textit{i.e.} a subject can only test one display. In this experiment, subjects are asked to successively retrieve 13 target sounds among the whole dataset. The target sounds are selected such that there are at least two target sounds in each top-level class of the semantic hierarchical structure of the dataset. To reduce order effects, target sounds are presented to each subject in a random order.

@@

60 subjects have completed the experiment for the urban sound dataset, 20 for each interface. 48 subjects have completed the experiment for the urban sound dataset: 17 for the PSD interface, 16 for the FSD interface and 15 for the AD interface.

@@

Outlier detection is an important step of any crowdsourcing experiment as experimenters do not control the environment in which the subjects perform the experiment \cite{komarov2013crowdsourcing,buchholz2011crowdsourcing}. A commonly used method to detect outliers in  human-computer interaction studies is to consider as outlier an observation which deviates of at least $\pm2$ standard deviation from the average \cite{komarov2013crowdsourcing}. However, as this latter method is not robust to the presence of isolated extreme observations (as it is often the case for crowdsourcing experiment), we follow the  method proposed by \cite{komarov2013crowdsourcing} and use the Inter-Quartile Range ($IQR$). With this approach, an observation is considered to be an outlier if its value is more than $3*IQR$ higher than the third quartile or more than $3*IQR$ less than the first quartile. This methods is applied in this study to the four following criteria :

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{gfx/urbanMusic1.eps} 
\end{center}
\caption{\label{urbanMusic} Average and standard deviation for (a,d) the total search durations, (b,e) the number of heard sounds and (c,f) the number of unique heard sounds; considering respectively the urban dataset (a,b,c) and the music dataset (d,e,f).}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{gfx/analyse_music2.eps} 
\end{center}
\caption{\label{fig2Music} Evolution of (a) the average search durations, (b) the average numbers of heard sounds, and (c) the average numbers of unique heard sounds at each target sound search, considering the music dataset. Lines are linear regression fits.}
\end{figure}

@@

\begin{itemize}
\item the \textit{total search duration},
\item the number of heard sounds during the whole experiment,
\item the number of unique sounds heard during the whole experiment. By "unique" we mean that, if a same sound is heard 10 times during the 13 searches of the  experiment, it counts only for one. 
\end{itemize}

@@

According to those results, a spatial display following a hierarchical organization of the dataset based on semantic values (PSD and FSD) is more efficient than an organization based on acoustic descriptors (AD) as it allows the users to quickly apprehend the dataset range and thus to specify their search. Furthermore the FSD interface allows the users to retrieve the 13 target sounds by listening to a smaller amount of sounds than the AD. However, those two effects are significantly compromised when users have to parse the entire hierarchy to reach the first target sound, as in the PSD display. It seems that enforcing an explicit exploration of the hierarchy disturbs or confuses the user instead of allowing him to learn the semantically motivated spatial organization of the classes.

@@

\section{Results on Musical Sounds} \label{resultsm}

@@

2 subjects are detected as outliers for the music experiment and removed from the analysis. 1 subject used PSD and 1 subject used FSD. The first outlier is detected by considering the \textit{absolute duration} (5 hours to complete the experiment) and the second by looking at the \textit{absolute duration} (1 hour and 30 minutes) as well as the \textit{total search duration} (1 hour and 15 minutes) and the total \textit{number of heard sounds} (901).

@@

PSD and FSD significantly reduce the search durations compared to an acoustic organization (AD), which was not the case for the urban sound dataset. This may be explained by the fact that the musical corpus has perhaps a more obvious semantic organization.

@@

Figure~\ref{fig2Music} shows the evolution of the metrics with respect to the target sound index. Results for the music dataset are similar to those observed for the urban dataset. An overall increase of efficiency with respect to the index of the search is observed for the tree interfaces. For the duration and the number of heard sounds (Figure~\ref{fig2Music}.a and .b) the FSD interface exhibits better performance. Considering the number of unique heard sounds (Figure~\ref{fig2Music}.c), the learning of the dataset range seems to be quicker for the FSD interface than for the PSD interface. Once again, results show that enforcing the parsing of the hierarchy is not beneficial.

>>>>

We thus consider in this paper three means of displaying sounds without relying on any textual representation, all based on a spatial organization of sounds. The first one, considered as a reference baseline, does not rely on any semantic information and positions sounds according to their acoustical properties (time averaged spectral features). The second and third displays exhibit a spatial organization based  on a predefined hierarchical semantic organization; those last two displays differ by how semantics are used.

@@

As a reference, an Acoustic Display (AD) provides a spatial representation based on acoustic descriptors. Each sound is described by time-averaged Mel-Frequency Cepstrum Coefficients (MFCCs) computed with standard parameter settings (13 lowest quefrency coefficients are kept out of 40). The Euclidean distance between time averaged features is then computed for each sound pair. A non metric multidimensional scaling (MDS) with Kruskal's normalized stress \cite{kruskal1964multidimensional} is then employed to project the data into two dimensions according to this distance, as shown on Figure~\ref{figXP3music}. In the case of instruments sounds, it exhibit a clear distinction between percussive (left) and sustained sounds (right).

@@

Among the several alternatives that exists, MFCCs features are chosen as they are widely used in the audio processing community. MDS is selected for 2D projection as it has a lower sensitivity to outliers than the PCA and does not rely on any parameter optimization, unlike the SOM approach.

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{gfx/music_mds.png} 
\end{center}
\caption{\label{figXP3music} Acoustical Display for the instrument sound dataset (AD). Colors and legend are removed during the experiment.}
\end{figure}

Two semantically oriented displays are then proposed, called Full Semantic Display (FSD) and Progressive Semantic Display (PSD). Both displays consider the hierarchical structure of the dataset to organize sounds. Each sound class is represented by a filled circle. Circles are then packed together according to the hierarchical semantic organization of the dataset, as shown in Figure~\ref{figSSF}. Thus, subclasses belonging to the same class are close to each others. Circle packing functions of the \emph{D3.js} (Data-Driven Documents) javascript library \cite{2011-d3} are used to distribute the sound classes in the space. Depending on the display, the user can either access each leaf class directly (FSD), or has to click through intermediate levels of the hierarchy (PSD).

@@

During this experiment, the three displays are compared using two corpora: the urban sound corpus and the instrument sound corpus. By comparing AD and xSD (the two semantic displays), the goal is to evaluate the  relative efficiencies of semantic based and acoustic based spatial configurations. By comparing FSD and PSD, we study the impact of an enforced hierarchical exploration of the corpus by the user. The goal is to check if constraining the user to first browse the high levels of the hierarchy helps him to grasp and memorize the spatial configuration and the organization of the sound classes.

@@

We evaluate and compare those three displays with a crowdsourcing web-based experiment\footnote{The tests are available at \url{http://soundthings.org/research/speedSoundFinding/index.html} and \url{http://soundthings.org/research/speedSoundFindingMusic/index.html}, each of the 3 displays is also available by appending a digit from 1 to 3 to \textit{index} in the above cited url.} for which we adopt a between subject approach, \textit{i.e.} a subject can only test one display. In this experiment, subjects are asked to successively retrieve 13 target sounds among the whole dataset. The target sounds are selected so that there are at least two target sounds in each top-level class of the semantic hierarchical structure of the dataset. To reduce order effects, target sounds are presented to each subject in a random order.

@@

60 subjects have completed the experiment for the urban sounds dataset, 20 for each interface. 48 subjects have completed the experiment for the instrument sounds dataset: 17 for the PSD interface, 16 for the FSD interface and 15 for the AD interface.

@@

Outlier detection is an important step of any crowdsourcing experiment as experimenters do not control the environment in which the subjects perform the experiment \cite{komarov2013crowdsourcing,buchholz2011crowdsourcing}. A commonly used method to detect outliers in  human-computer interaction studies is to consider as outlier an observation which deviates of at least $\pm2$ standard deviations from the average \cite{komarov2013crowdsourcing}. However, as this latter method is not robust to the presence of isolated extreme observations (as it is often the case for crowdsourcing experiment), we follow the  method proposed by \cite{komarov2013crowdsourcing} and use the Inter-Quartile Range ($IQR$). With this approach, an observation is considered to be an outlier if its value is more than $3*IQR$ higher than the third quartile or more than $3*IQR$ less than the first quartile. This methods is applied in this study to the four following criteria :

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{gfx/urbanMusic1.eps} 
\end{center}
\caption{\label{urbanMusic} Average and standard deviation for (a,d) the total search durations, (b,e) the number of heard sounds and (c,f) the number of unique heard sounds; considering respectively the urban dataset (a,b,c) and the instrument dataset (d,e,f).}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{gfx/analyse_music2.eps} 
\end{center}
\caption{\label{fig2Music} Evolution of (a) the average search durations, (b) the average numbers of heard sounds, and (c) the average numbers of unique heard sounds at each target sound search, considering the instrument dataset. Lines are linear regression fits.}
\end{figure}

@@

\begin{itemize}
\item the \textit{total search duration},
\item the number of heard sounds during the whole experiment,
\item the number of unique sounds heard during the whole experiment. By ``,unique\,'' we mean that, if a same sound is heard 10 times during the 13 searches of the  experiment, it is only counted once. 
\end{itemize}

@@

According to those results, a spatial display following a hierarchical organization of the dataset based on semantic values (PSD and FSD) is more efficient than an organization based on acoustic descriptors (AD) as it allows the users to quickly apprehend the dataset range and thus to specify their search. Furthermore the FSD interface allows the users to retrieve the 13 target sounds by listening to a smaller amount of sounds than the AD. However, those two effects are significantly compromised when users have to parse the entire hierarchy to reach the first target sound, as in the PSD display. It seems that enforcing an explicit exploration of the hierarchy disturbs or confuses the user instead of allowing them to learn the semantically motivated spatial organization of the classes.

@@

\section{Results on Instrumental Sounds} \label{resultsm}

@@

2 subjects are detected as outliers for the instrument sounds experiment and removed from the analysis. 1 subject used PSD and 1 subject used FSD. The first outlier is detected by considering the \textit{absolute duration} (5 hours to complete the experiment) and the second by looking at the \textit{absolute duration} (1 hour and 30 minutes) as well as the \textit{total search duration} (1 hour and 15 minutes) and the total \textit{number of heard sounds} (901).

@@

PSD and FSD significantly reduce the search durations compared to an acoustic organization (AD), which was not the case for the urban sound dataset. This may be explained by the fact that the instrumental corpus has perhaps a more obvious semantic organization.

@@

Figure~\ref{fig2Music} shows the evolution of the metrics with respect to the target sound index. Results for the instrument dataset are similar to those observed for the urban dataset. An overall increase of efficiency with respect to the index of the search is observed for the tree interfaces. For the duration and the number of heard sounds (Figure~\ref{fig2Music}.a and .b) the FSD interface exhibits better performance. Considering the number of unique heard sounds (Figure~\ref{fig2Music}.c), the learning of the dataset range seems to be quicker for the FSD interface than for the PSD interface. Once again, results show that enforcing the parsing of the hierarchy is not beneficial.
